### 2016-07-25 latex/input/abstract.tex
completed fixes of gicf with mini batch size randomization, proper dot products, top k instances, randomized weight initialization; added temporary latex template for paper

%abs

We demonstrate a computational method to predict the clinical
phenotypes of a patient from raw metagenomic sequence read data. We  compared 
two state of the art programs for annotating the sequence data, 
UCLUST and Kraken, and using their output for feature generation. 
We apply these programs 
to a set of over 1.3 million reads from 
904 patients, some of whom have liver cirrhosis, encephalopathy due to liver cirrhosis, or 
neither disease. Once the reads have been processed by UCLUST or 
Kraken, we use  Support Vector Machines to setup the 
clinical phenotype prediction problem.  
%
We find that too many false negatives are being predicted by 
the classifier. In order to address the issue, we scale features to
improve the classification model and evaluate the 
end results on a held-out test set. We
demonstrate our approach works quickly and 
accurately with an 85.64\% success rate when we use the 
UCLUST representation. We also find that UCLUST generally performs better than Kraken, with the latter having an 80.66\% success rate. We also test our classifier on several subsets of the data, with success rates ranging from 69.81\% to 96.72\%.