### 2016-07-25 latex/input/intro.tex

%Intro

The human body is one of the 
most densely populated microbial 
environments in the world, with human host cells interacting 
with  more than $10^{14}$ microbial cells. Collectively, the human 
and microbial cells are referred to as the 
\emph{human microbiome} \cite{turnbaugh_human_2007,backhed_host-bacterial_2005}. 
%
%
%
Biotechnological  advances 
can 
interrogate the microbial 
communities present across three primary dimensions: (i) clinically different
patients like those suffering from liver disease versus healthy controls (ii) different
patient sites and tissues like skin and the gastrointestinal tract 
and (iii) across different patient visits~\cite{Costello12182009,Qin:2010fk} and before/after treatment. 
%
Methods to query this  microbial
data include DNA sequencing of 
specific marker genes such as the 16S rRNA marker genes~\cite{Woese97}, 
sequencing
the entire pool of microbial genomes at once (metagenomics), or 
sequencing the mixture of 
expressed gene transcripts (metatranscriptomes).
%
These sequencing methods along 
with proteomics and mass spectrometry
allow us to study the
biotransformations caused by the microbial communities. 
%
Several researchers and clinicians have embarked on
studies of the pathogenic and clinical 
role played by the microbiome with respect to
human health and disease conditions.

In this study, we specifically developed
a computational method to predict the clinical phenotype
of a patient i.e., a pipeline to 
predict whether or not the patient has a specific type and level of  a disease.  This 
pipeline
is able to make the phenotypic
predictions from input metagenomic sequences using post-processing 
with unsupervised binning approaches (clustering) or  supervised search based 
taxa profiling approaches.  This method is designed to 
run quickly and produce accurate and sensitive 
results. We  compare and evaluate two state-of-the-art 
approaches, one for binning (UCLUST \cite{Edgar10}) and one for taxonomic profiling (KRAKEN \cite{Wood14}) 
to represent the input metagenome sequences as features for classification. The classification
formulation uses a binary support vector machine based classifier \cite{vap95}. We also 
evaluate the performance of preliminary 
feature scaling and engineering  for classification purposes.

We specifically evaluated our approach on 
microbiome samples of patients suffering from hepatic encephalopathy due 
to liver cirrhosis. 
%
We found that 
UCLUST performed better than Kraken on our data 
set, although Kraken was faster. We were also able to 
engineer the data to improve the accuracy of the classifier. Our 
pipeline had 85.64\% accuracy using UCLUST and 80.66\% accuracy 
using Kraken.

%For our data set with 1.3 million reads, UCLUST took 
%180 seconds to run and Kraken took 120 seconds. 
%