### 2016-08-04 latex/input/materials.tex

\subsection{Dataset Description}

We used data from a well-known Metagenome-Wide Association Study by Qin et al. of Type 2 Diabetes (T2D) in Chinese patients \cite{qin041012}. The total dataset used in this study contains 368 patients \cite{qin041012}, but only 218 of them were labeled and available online. Each patient file was downloaded from NCBI\footnote{http://trace.ncbi.nlm.nih.gov/Traces/study/?acc=SRP011011}
and converted to FASTQ format using the SRA toolkit\footnote{http://www.ncbi.nlm.nih.gov/Traces/sra/sra.cgi?view=software}.
The labels were found on the EBI website\footnote{https://www.ebi.ac.uk/metagenomics/projects/SRP011011}. The total size of these 218 FASTQ files was 2.3 terabytes, with an average size of 10.82 gigabytes per patient file. Qin et al. develop a simple T2D classifier using a minimum redundancy–maximum relevance (mRMR) method \cite{peng05} for feature selection and an SVM for classification, based on the R packages "sideChannelAttack" and "1071", respectively \cite{qin041012}. They achieved an AUC-ROC of 0.81 by training a classifier on 345 of the patients and using the remaining 23 as a test set \cite{qin041012}. While we did not have access to all of the same patients that they had, we were able to replicate their method and test it on the subset of data that was publicly available. The study also called for more extensive testing of gut microbiota classifiers \cite{qin041012}.

\subsection{Software and Hardware Details}
We used the ARGO computing cluster available at George Mason University\footnote{http://orc.gmu.edu/research-computing/argo-cluster/argo-hardware-specs/}. The clustering and classification phases were run on one of the compute nodes available on the cluster. The cluster is configured with 35 Dell C8220 Compute Nodes, each with dual Intel Xeon E5-2670 (2.60GHz) 8 core CPUs, with 64 GB RAM. (Total Cores 528 and 1056 total threads, RAM$>$2TB). Source codes for 
SOAPdenovo2\footnote{SOAPdenovo2: http://soap.genomics.org.cn/soapdenovo.html} \cite{luo12}, UCLUST\footnote{UCLUST: http://www.drive5.com/uclust/downloads1\_{}2\_{}22q.html}    \cite{Edgar10}, and svm-light\footnote{svm-light: http://svmlight.joachims.org/} \cite{joachims08}
their respective websites and compiled on the Argo platform.