### 2016-08-23 latex/input/methods.tex

%methods

\begin{figure*}[t]
\centering
\includegraphics[scale=0.5]{./mil-metagenomics-pipeline.png}
\caption{This diagram illustrates the entire CAMIL pipeline. Patient files with FASTQ metagenomic reads are individually assembled using SOAPdenovo2, then combined into one file and clustered with UCLUST. We extract features according to either the D-BoW or H-BoW method, and classify the patients using the feature vectors with svm-light.} \label{pipeline}
\end{figure*}

\subsection{Overview}

Our proposed pipeline involves a number of steps, which serve a variety of purposes. For each patient file, we assembled the sequence reads, which served the dual purpose of generating larger contigs that contain more functional biological information and reducing the dataset size by discarding reads that could not be assembled. The clustering step assigns the contigs to certain clusters, which represent functionally similar microbes, and thus establish classes of instances that can be used as features for the classifier. We  applied our feature extraction method, which uses either the Histogram-based or Distance-based BoW methods. Finally, we used an SVM-based classifier to predict patient phenotype, and used several metrics to assess its accuracy. We used the SVM's decision boundary to infer information about which clusters of instances were most or least indicative of the phenotype, discussed further in the Results section. Aside from the patient labels, this process is entirely de novo, and does not consult any external databases. An illustration of the pipeline is shown in Figure \ref{pipeline} on page \pageref{pipeline}.

\subsection{Assembly with SOAPdenovo2}

For our assembly step, we used SOAPdenovo2 \cite{luo12}, which is a de novo metagenome assembly tool. This tool was also used for assembly in the MGWAS study that our data came from. We use this method in order to compare ourselves with the MGWAS study, and because Soapdenovo2 has been proven to be one of the fastest assemblers \cite{peng12}. We tested a number of different combinations of parameters, and found that the best results came when we cut reads off after 100 base pairs (reads were 180 base pairs long originally) and used a k-mer size of 51. The average insert size was set to 350, in accordance with the reported average insert size from the MGWAS study that we used data from \cite{qin041012}. The patient files needed to be assembled separately, in order to avoid assembling reads from different patients together. Conversely, all contigs need to be in one file for clustering, to avoid inconsistent cluster assignments between different patients. Thus, we combined the contigs from each assembled patient file into a single for clustering.

\subsection{Clustering with UCLUST}

We used UCLUST \cite{Edgar10}, which has been shown to be among the simplest, fastest, and most effective clustering methods \cite{bonder090112, sun042711}. Since our contigs were not ordered, we used the usersort option, and we set the sequence match threshold to 40\%, which means that two reads needed to have 40\% of the same nucleotides to be in the same cluster. For instance, between two strings of length 100, at least 40 places in each of those strings would have to contain the same nucleotide (represented as A, T, G, or C). New contigs that did not match at least 40\% to any of the existing contigs would form the seed of a new cluster. This value of 40\% was found to be the best value based on our experiments with this dataset. Other values tried, including 50\%, 75\%, and 90\%, led to many clusters with very few reads per cluster.

\subsection{Feature Extraction}

We test both the Distance-based and Histogram-based feature extraction methods within the CAMIL pipeline. For each patient, we initialized a vector with length equal to the number of clusters. The value for each scalar in that patient's feature vector would then represent either the read that matched most closely to that cluster centroid (Distance-based), or the number of reads from that cluster that were present (Histogram-based. An example of this is illustrated as a part of Figure \ref{pipeline}. We implemented this feature extraction method in Python.

\subsection{Classification and Evaluation Metrics}

We performed classification with a standard SVM classifier using the generated feature vectors; in this case, we used svm-light \cite{joachims08}. The choice of classifier is not very important for D-BoW methods \cite{amores13}.

We can assess the success of our classifier in several ways. The simplest measure, accuracy, measures the percentage of instances 
that are classified correctly, represented by 
\begin{equation}
Accuracy = (TP + TN)/ (TP + TN + FP + FN)  \label{eqn:acc} 
\end{equation}
where TP, TN, FP and FN represents true positives, true negatives, false positives and false negatives respectively.

Accuracy as an evaluation metric can be biased if one of the classes 
(positive or negative)  has a larger number of examples than the other.   
Precision measures the percentage of positive predictions that 
were correct, 
whereas recall measures the percentage of positive 
examples that were correctly predicted (or retrieved). 
We can represent Precision and Recall as:

\begin{equation}
Precision = TP / (TP + FP). \label{eqn:prec}
\end{equation}
\begin{equation}
Recall = TP / (TP +FN). \label{eqn:roc}
\end{equation}

The F1 score captures the trade-offs between precision and recall in a
single metric and is the harmonic mean of precision and recall, given by:
\begin{equation}
F1-Score = 2 * (Precision * Recall)/ (Precision + Recall). \label{eqn:f1}
\end{equation}

Finally, we also use the Area Under Curve of the Receiver Operating Characteristic (AUC-ROC), which measures the performance of the classifier as the decision boundary threshold is moved. The SVM classifier generally predicts a group label to be negative if the predicted label for that group was less than 0 and predicts a group label to be positive otherwise. The AUC-ROC measures the performance of the classifier as the threshold is varied to more or less than 0. In effect, it measures how far off incorrect predictions were from being correct. AUC-ROC plots True Positive Rate (TPR) versus False Positive Rate (FPR), given by:
\begin{equation}
TPR  = TP / (TP + FN). \label{eqn:roc}
\end{equation}
\begin{equation}
FPR = FP / (FP + TN). \label{eqn:prec}
\end{equation}

\subsection{Deriving Instance ``Labels"}

\begin{figure}[t]
\centering
\includegraphics[scale=0.5]{./instance-labels.png}
\caption{This diagram illustrates why static instance labels are not sufficient for phenotype prediction. A patient with 6 of the blue microbe or 6 of the green microbe may be healthy, while a patient with 3 of each is sick. Static instance labels cannot capture this relationship. This is also explained by Amores in his MIL taxonomy \cite{amores13}.} \label{instance-labels}
\end{figure}

One of the benefits of using Multiple Instance Learning methods is that we can attempt to discover instance ``labels". In fact, we did not attempt to apply static, unchanging labels to individual reads or clusters, since organisms are affected by their interactions with each other. For instance, a patient with X amount of microbe A or X amount of microbe B may be healthy, but with X/2 amount of microbe A and X/2 amount of microbe B they may be sick. This simple example is illustrated in Figure \ref{instance-labels}. We can infer from the SVM decision boundary which clusters appear to be most relevant to the disease diagnosis. Since feature vectors are multiplied by the weight vector of the decision boundary to determine the label of the patient, we can assume that clusters with the highest weights in the weight vector are most relevant to the disease diagnosis. For instance, if the \emph{i}th scalar in the weight vector is has the highest value of any of the weights, then cluster \emph{i} is likely to play a major role in the disease pathology. Similarly, the most negative weights in the weight vector indicate clusters whose presence in a patient indicates that they likely do not have the disease. Because the data is metagenomic, the clusters represent both phylogenetic and functional similarity, so identifying the most relevant clusters can help discover more about the pathology of the disease. For Type 2 Diabetes, which is a complex phenotype and a disease that is both common and deadly, this is potentially quite valuable.