### 2016-10-24 latex/input/conclusion.tex
initial paper updated further explaining feature extraction method and reducing explanation of assembly

%Conclusion

We have demonstrated an effective and efficient computational pipeline for classifying patient phenotype based on metagenomic data. We have demonstrated that even relatively simple de novo assembly and clustering methods, when used within this pipeline, lead to significantly better performance results than the standard classifier used in the original Metagenome-Wide Association Study and other common MIL methods. We would like to emphasize that, while we used a particular type 2 diabetes dataset in this paper, the methods described here are general-purpose and could be used for any metagenomic dataset. We have shown how to infer the most important OTUs in the disease pathology by using the SVM decision boundary and discussed the clinical importance of this ability. More generally, we have shown the effectiveness of Multiple Instance Learning methods within metagenomics and phenotype prediction, particularly Bag of Words methods. CAMIL is a relatively simple and easy to implement pipeline that has both shown strong results and significant potential for even further improvement. Future work could revolve around improving individual parts of the pipeline, such as using better assembly and clustering methods, application of different multiple instance learning methods (other than Bag of Words), and further attempts to generate more specific instance level or cluster level information and validate that information against the known pathology of various diseases.